Sound and Automated Synthesis of Digital Controllers for Physical Plants

### 2016-10-03 paper/paper.tex
hscc17
Signed-off-by: Alessandro Abate <aabate@Alessandros-MacBook-Pro.local>

\blue{Please insert a figure clearly distinguishing between continuous and discretised/quantised plant - a template is in a picture sent out by Lucas late last week. 
We should have 
\begin{itemize}
\item 
$G(s) = \frac{s^{M_G}b_{0}+s^{M_G-1}b_{1}+...+b_{M_G}}{s^{N_G}a_{0}+s^{N_G-1}a_{1}+...+a_{N_G}}$ -- this is the physical plant, which is focus of interest and to which we refer our specifications/properties 
\item 
$G(z, T) = \frac{\bar b_{0}+\bar b_{1}z^{-1}+...+\bar b_{M_G}z^{-M_G}}{\bar a_{0}+\bar a_{1}z^{-1}+...+\bar a_{N_G}z^{-N_G}}$ -- this is the time-discretised plant (notice the notation encompasses explicitly the sample time $T$; please write out explicitly the relationship between the $a_i$ and the $\bar a_i$ and likewise for the numerator's coefficients 
(in particular, are the orders $M_G, N_G$ the same? 
\item 
we have $Y(z) = G(z,T) U(z)$; then, we introduce $\hat Y(z) \doteq Y(z) + \nu (z)$, so that 
$$
\frac{\hat Y(z)}{U(z)} = G(z,T) + \frac{\nu (z)}{U(z)} = G(z,T) + \Delta G_Q(z) \doteq \hat G(z) ,  
$$ 
where $\Delta G_Q(z)$ is an error term accounting for the quantisation effects 
\item 
finally, we encompass the errors resulting from employing finite parameters representation, 
which we encompass in a term $\Delta G_P(z)$: is this error additive, so that we would get to a 
$$
\tilde G(z) \doteq \hat G(z) + \Delta G_P(z) = G(z,T) + \Delta G_Q(z) + \Delta G_P(z)?  
$$ 
Dario, can you write it out explicitly? In other words, if 
$$
\hat G(z) = G(z,T) + \Delta G_Q(z) = \frac{\hat b_{0}+\hat b_{1}z^{-1}+...+\hat b_{M_G}z^{-M_G}}{\hat a_{0}+\hat a_{1}z^{-1}+...+\hat a_{N_G}z^{-N_G}},  
$$ 
then what is the explicit expression of 
$$
\tilde G(z) = \frac{\tilde b_{0}+\tilde b_{1}z^{-1}+...+\tilde b_{M_G}z^{-M_G}}{\tilde a_{0}+\tilde a_{1}z^{-1}+...+\tilde a_{N_G}z^{-N_G}}?
$$
(Notice that the orders $M_G, N_G$ need to be properly taken care of.)
\item 
Now, it's over the transfer function $\tilde G(z)$ that we want to synthesise the parameters of a template $C(z)$. 
This controller's transfer function will be subject to the same quantisation and precision limits as the plant's: 
we need to argue that the way we do synthesis affects manipulation of quantities that relate to specific errors and be explicit about these. 
\item[]
\item 
An additional source of error can be related to the fact that a plant transfer function, 
per se a physical entity, is actually represented in a program imprecisely: 
this would mean that $G(s)$ is actually 
$$
G^\ast (s) \doteq \frac{s^{M_G^\ast}b_{0}+s^{M_G^\ast-1}b_{1}+...+b_{M_G}}{s^{N_G^\ast}a_{0}+s^{N_G^\ast-1}a_{1}+...+a_{N_G}}, 
$$ 
which perhaps can indeed be re-written as $G^\ast (s) = G(s) + \Delta G_P(s)$? (This proof should come from that above.) 
We should argue that this error is encompassed by the one derived above (or isn't it?). 
\end{itemize}
}

@@

In this study, the methodology used for verifying closed-loop digital
control system is based on the Digital-System Verifier
(DSVerifier)~\cite{IsmailBCFF15}, which checks the stability of control
systems considering finite-word length (FWL) effects in the digital
controller and uncertainty parameters in the plant model (plant
intervals)~\cite{Bessa16}.  Let $C(z)$ be a digital controller and $G(z)$ be
a plant model given as
%
\begin{align}
\small
\label{controller_plant_tf}
C(z)&=\frac{\beta_{0}+\beta_{1}z^{-1}+...+\beta_{M_C}z^{-M_C}}{\alpha_{0}+\alpha_{1}z^{-1}+...+\alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
G(z)&=\frac{b_{0}+b_{1}z^{-1}+...+b_{M_G}z^{-M_G}}{a_{0}+a_{1}z^{-1}+...+a_{N_G}z^{-N_G}}.
\end{align}
%
\noindent where $\beta$ and $\alpha$ are the controller's coefficients, $b$
and $a$ are the plant's coefficients, and $N$ and $M$ are the number of bits
of the integer and fractional part, respectively.  Note that $G(z)$ is a
discrete model of a real (and continuous) plant, {\it i.e.}, a continuous
model of the plant must be discretized to obtain the discrete equivalent
coefficients $b$ and $a$~\cite{Astrom08}.  Although there are several
discretization methods available in the literature~\cite{Franklin15}, the
zero-order hold (ZOH) discretization is more suitable for representing the
sampling processes that are typically employed by hybrid
systems~\cite{istepanian2012digital}.

Let $C(z)$ and $G(z)$ transfer functions be arranged in vectors $c_0$ and
$g_0$, respectively.  For each $C(z)$ implementation, there is a function
$\mathit{FWL}[\cdot]:\mathbb{R}^{N_C+M_C}\rightarrow
Q[\mathbb{R}^{N_C+M_C}]$, which applies FWL effects to the controller
$C(z)$, where $Q[\mathbb{R}]$ represents the quantized set of representable
real numbers in the chosen implementation format \param{N_C}{M_C}.  The
quantized controller parameters vector ($c$) is generated via
$\mathit{FWL}[c_0]$.  The uncertain plant parameters ($g$) can be expressed
as
%
\begin{equation}
g = g_0 + \Delta{g} = \left [
				\begin{tabular}{c}
				$b_{0}+\Delta{b_{0}} ~ b_{1}+\Delta{b_{1}} ~ ... ~ b_{M_G}+\Delta{b_{M_G}}$ \\%~ 
				$a_{0}+\Delta{a_{0}} ~ a_{1}+\Delta{a_{1}} ~ ... ~ a_{N_G}+\Delta{a_{N_G}}$
				\end{tabular}
				\right ],
\end{equation}

@@

@@

\noindent where $\Delta{G_{Q}(z)}$ represents the quantization uncertainty
model and $\Delta{G_{P}(z)}$ describes the parametric uncertainty model. 
Thus, a model for $\Delta{G_{Q}(z)}$ that can represent the quantization
noise effect in the plant output, and consequently the closed-loop system
output, is proposed.  For convenience, the parametric uncertainties will be
despised here.

@@

First, some assumptions have to be established to obtain a model for
$\Delta{G_{Q}(z)}$.
%
\begin{myassumption}
\label{whitenoise}
%
The quantization noises $\nu_{1}$ (from the $Q1$) and $\nu_{2}$ (from the
$Q2$) are white noises with zero mean.
%
\end{myassumption}

@@

\begin{myassumption}
%
The controller and plant are linear.  In particular, the digital controller
does not contain nonlinearities such as overflows, saturation, and
limit-cycle oscillations.
%
\end{myassumption}

\begin{myassumption}
%
The DAC interpolator is an ideal ZOH; it takes the discrete value of control
signal $u(k)$ at instant $k$ and produces a rectangular pulse of height
$u(k)$ and width equal to the sample time $T$.
%
\end{myassumption}

Given those assumptions, the present approach is further based on the following lemmas.
%
\begin{mylemma}
%
\cite{widrow1956} The second moment (variance) of the PDF of quantization
noise is given by \textcolor{red}{expand the acronym PDF}
%
\begin{equation}
\label{eq:variancelemma}
\sigma_{Q}^{2}=\frac{q^{2}}{12},
\end{equation}
\end{mylemma}

@@

\begin{mylemma} \cite{astrom1997computer}
%
The pulse transfer function $G(z)$ is equivalent at the sample times to the
continuous transfer function $G(s)$ driven by ZOH.  This pulse transfer
function can be computed as
%
\begin{equation}
\label{eq:pulsetf}
G(z)=(1-z^{-1})\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\frac{G(s)}{s}}\right\rbrace_{t=kT}}\right\rbrace.
\end{equation}
\end{mylemma}

@@

\noindent where $\sigma_{Q}^{2}$ is the variance of quantization noise and
$U(z)$ is the z-transform of control signal ({\it i.e.}, the z-transform of
the signal from the digital controller), which is provided to the plant by
means of ZOH.

@@

Eq.~\eqref{eq:deltag_var} indicates that the modeling of quantization noise
effect on the plant transfer function demands an estimation of the variance
of quantization noise $\sigma_{Q}^{2}$.  To obtain such estimation, the
methodology presented by Widrow and Koll\'ar~\cite{widrow2008quantization}
is used.  The authors provide an algorithm for computing the mean square of
the output noise due to $Q1$ and $Q2$.  Considering that each quantizer
contribution can be represented by a white noise with zero mean
(Assumption~\ref{whitenoise}), the mean square can be considered as the
variance of noise ($\sigma_{Q}^{2}$), which can be calculated as
%
\begin{equation}
\sigma_{Q}^{2}=\frac{\lambda_{1}^{2}\cdot \sum {h_{Q1}^{2}}+\lambda_{2}^{2}\cdot \sum {h_{Q2}^{2}}}{12},
\end{equation}
%
\noindent where $\lambda_{1}$ and $\lambda_{2}$ are the scale factor for the
quantization $Q1$ and $Q2$, respectively; and $\sum {h_{Q1}^{2}}$ and $\sum
{h_{Q2}^{2}}$ are the squares sums of the impulse response obtained by means
of a transfer function from $Q1$ and $Q2$ to the plant output, respectively.

@@

These equations are sufficient for computing $\sigma_{Q}^{2}$.  However,
there is still a gap for a complete model for $\Delta{G_{Q}(z)}$ in
Eq.~\eqref{eq:deltag_var}, {\it i.e.}, the z-transform of control action
$U(z)$.  This indicates that the quantization noise effect on a discrete
plant model mainly depends on three main factors:
%
\begin{itemize}
%
\item \textbf{Sample time:} the sample time effect is reflected over pulse transfer function that covers the effects of DAC ant its interpolator (ZOH);
%
\item \textbf{ADC and DAC step: } these are the main component of quantization noise power, but an $8$-bits converter already reduces the effect of quantization noises in the output for negligible values.
%
\item \textbf{Signal-to-noise ratio (SNR): } the dependence on the value of $U(z)$ shows the importance of the SNR for handling quantization noise; for a high SNR of the digital controller, the quantization effect is negligible. This happens during the transient of the system, when high control actions are employed. However, it can be relevant during the steady state. A digital controller with integral action might be sufficient for eliminating such influence on the steady state.
%
\end{itemize}

@@

A more confident description of $\Delta{G_{Q}(z)}$ can be obtained
considering FWL effects on digital controller coefficients, as described in
Bessa {\it et al.}~\cite{Bessa16}, {\it i.e.}, using $FWL[C(z)]$ instead of
the nominal $C(z)$:
%
\begin{equation}
\label{eq:deltaG_final}
\Delta{G_{Q}(z)}=\frac{\sigma^{2}_{Q}}{r}\cdot \frac{1+FWL[C(z)]G(z)}{FWL[C(z)]} \cdot \frac{z-1}{z}.
\end{equation}

>>>>

\blue{Please insert a figure clearly distinguishing between continuous and discretised/quantised plant - a template is in a picture sent out by Lucas late last week. Please introduced signals on the plot.

Let us straighten the notations next. 
We should have 
\begin{itemize}
\item 
$G(s) = \frac{s^{M_G}b_{0}+s^{M_G-1}b_{1}+...+b_{M_G}}{s^{N_G}a_{0}+s^{N_G-1}a_{1}+...+a_{N_G}}$ -- this is the physical plant, which is focus of interest and to which we refer our specifications/properties 
\item 
$G(z, T) = \frac{\bar b_{0}+\bar b_{1}z^{-1}+...+\bar b_{M_G}z^{-M_G}}{\bar a_{0}+\bar a_{1}z^{-1}+...+\bar a_{N_G}z^{-N_G}}$ -- this is the time-discretised plant (notice the notation encompasses explicitly the sample time $T$; please write out explicitly the relationship between the $a_i$ and the $\bar a_i$ and likewise for the numerator's coefficients 
(in particular, are the orders $M_G, N_G$ the same? 
\item 
we have $Y(z) = G(z,T) U(z)$; then, we introduce $\hat Y(z) \doteq Y(z) + \nu (z)$, so that 
$$
\frac{\hat Y(z)}{U(z)} = G(z,T) + \frac{\nu (z)}{U(z)} = G(z,T) + \Delta G_Q(z) \doteq \hat G(z) ,  
$$ 
where $\Delta G_Q(z)$ is an error term accounting for the quantisation effects 
\item 
finally, we encompass the errors resulting from employing finite parameters representation, 
which we encompass in a term $\Delta G_P(z)$: is this error additive, so that we would get to a 
$$
\tilde G(z) \doteq \hat G(z) + \Delta G_P(z) = G(z,T) + \Delta G_Q(z) + \Delta G_P(z)?  
$$ 
Dario, can you write it out explicitly? In other words, if 
$$
\hat G(z) = G(z,T) + \Delta G_Q(z) = \frac{\hat b_{0}+\hat b_{1}z^{-1}+...+\hat b_{M_G}z^{-M_G}}{\hat a_{0}+\hat a_{1}z^{-1}+...+\hat a_{N_G}z^{-N_G}},  
$$ 
then what is the explicit expression of 
$$
\tilde G(z) = \frac{\tilde b_{0}+\tilde b_{1}z^{-1}+...+\tilde b_{M_G}z^{-M_G}}{\tilde a_{0}+\tilde a_{1}z^{-1}+...+\tilde a_{N_G}z^{-N_G}}?
$$
(Notice that the orders $M_G, N_G$ need to be properly taken care of.)
\item 
Now, it's over the transfer function $\tilde G(z)$ that we want to synthesise the parameters of a template $C(z)$. 
This controller's transfer function will be subject to the same quantisation and precision limits as the plant's: 
we need to argue that the way we do synthesis affects manipulation of quantities that relate to specific errors and be explicit about these. 
\item[]
\item 
An additional source of error can be related to the fact that a plant transfer function, 
per se a physical entity, is actually represented in a program imprecisely: 
this would mean that $G(s)$ is actually 
$$
G^\ast (s) \doteq \frac{s^{M_G^\ast}b_{0}+s^{M_G^\ast-1}b_{1}+...+b_{M_G}}{s^{N_G^\ast}a_{0}+s^{N_G^\ast-1}a_{1}+...+a_{N_G}}, 
$$ 
which perhaps can indeed be re-written as $G^\ast (s) = G(s) + \Delta G_P(s)$? (This proof should come from that above.) 
We should argue that this error is encompassed by the one derived above (or isn't it?). 
\end{itemize}
}

@@

\red{[This section and the next should be merged into a crisp section with the background on DSVerifier, if not included in previous work? 
Please stick to the introduced notations. Please clarify what the function FWL does, compared to the general approximations introduced above.]}

In this study, the methodology used for verifying closed-loop digital
control system is based on the Digital-System Verifier
(DSVerifier)~\cite{IsmailBCFF15}, which checks the stability of control
systems considering finite-word length (FWL) effects in the digital
controller and uncertainty parameters in the plant model (plant
intervals)~\cite{Bessa16}.  Let $C(z)$ be a digital controller and $G(z)$ be
a plant model given as
%
\begin{align}
\small
\label{controller_plant_tf}
C(z)&=\frac{\beta_{0}+\beta_{1}z^{-1}+...+\beta_{M_C}z^{-M_C}}{\alpha_{0}+\alpha_{1}z^{-1}+...+\alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
G(z)&=\frac{b_{0}+b_{1}z^{-1}+...+b_{M_G}z^{-M_G}}{a_{0}+a_{1}z^{-1}+...+a_{N_G}z^{-N_G}}.
\end{align}
%
\noindent where $\beta$ and $\alpha$ are the controller's coefficients, $b$
and $a$ are the plant's coefficients, and $N$ and $M$ are the number of bits
of the integer and fractional part, respectively.  Note that $G(z)$ is a
discrete model of a real (and continuous) plant, {\it i.e.}, a continuous
model of the plant must be discretized to obtain the discrete equivalent
coefficients $b$ and $a$~\cite{Astrom08} \red{[pls say how]}.  
Although there are several
discretization methods available in the literature~\cite{Franklin15}, the
zero-order hold (ZOH) discretization is more suitable for representing the
sampling processes that are typically employed by hybrid
systems~\cite{istepanian2012digital} 
\red{[but the ZOH is for the DAC, not for the ADC?]}.

Let $C(z)$ and $G(z)$ transfer functions be arranged in vectors $c_0$ and
$g_0$, respectively.  For each $C(z)$ implementation, there is a function
$\mathit{FWL}[\cdot]:\mathbb{R}^{N_C+M_C}\rightarrow
Q[\mathbb{R}^{N_C+M_C}]$, which applies FWL effects to the controller
$C(z)$, where $Q[\mathbb{R}]$ represents the quantized set of representable
real numbers in the chosen implementation format \param{N_C}{M_C} \red{[clarify this notation]}.  The
quantized controller parameters vector ($c$) is generated via
$\mathit{FWL}[c_0]$.  The uncertain plant parameters ($g$) can be expressed
as
%
\begin{equation}
g = g_0 + \Delta{g} = \left [
				\begin{tabular}{c}
				$b_{0}+\Delta{b_{0}} ~ b_{1}+\Delta{b_{1}} ~ ... ~ b_{M_G}+\Delta{b_{M_G}}$ \\%~ 
				$a_{0}+\Delta{a_{0}} ~ a_{1}+\Delta{a_{1}} ~ ... ~ a_{N_G}+\Delta{a_{N_G}}$
				\end{tabular}
				\right ],
\end{equation}

@@

\red{[The following description can be made shorter, crisper?]}

@@

\noindent where $\Delta{G_{Q}(z)}$ represents the quantization uncertainty
model and $\Delta{G_{P}(z)}$ describes the parametric uncertainty model. 
Thus, a model for $\Delta{G_{Q}(z)}$ that can represent the quantization
noise effect in the plant output, and consequently the closed-loop system
output, is proposed.  For convenience, the parametric uncertainties will be
described here.

@@

First, some assumptions have to be established to obtain a model for $\Delta{G_{Q}(z)}$. 
\blue{[Here we need to be careful: we first argue that quantisation (spelled with an s, so that ER II is happy) effects can be represented as an additive noise (great), which however is a stochastic signal. But we handle the noise deterministically. How so? Where is the catch? ]}
%
\begin{myassumption}
\label{whitenoise}
%
The quantization noises $\nu_{1}$ (from the $Q1$) and $\nu_{2}$ (from the
$Q2$) \red{[do we need the second? See my further comments below.]} are white noises with zero mean.
%
\end{myassumption}

@@

\begin{myassumption}
%
The controller and plant are linear.  In particular, the digital controller
does not contain nonlinearities such as overflows, saturation, and
limit-cycle oscillations. \red{[explain difference btw overflows and saturations. Also, do we care about limit cycles? I'd quite happily disregard them.]}
%
\end{myassumption}

\begin{myassumption}
%
The DAC interpolator is an ideal ZOH; it takes the discrete value of control
signal $u(k)$ at instant $k$ and produces a rectangular pulse of height
$u(k)$ and width equal to the sample time $T$. \red{[Right: so under this assumption, we really only focus on the quantisation effects introduced by the ADC, correct? Let us not mix up the role and the influence of the two blocks.]}
%
\end{myassumption}

Given those assumptions, the present approach is further based on the following lemmas. 
\red{[So we have introduced two quantisation errors, properly indexed. Next we discuss a non-indexed entity - pls clarify or comment.]}
%
\begin{mylemma}
%
\cite{widrow1956} The second moment (variance) of the PDF of quantization
noise is given by \textcolor{red}{expand the acronym PDF}
%
\begin{equation}
\label{eq:variancelemma}
\sigma_{Q}^{2}=\frac{q^{2}}{12},
\end{equation}
\end{mylemma}

@@

\begin{mylemma} \cite{astrom1997computer}
%
The pulse transfer function \red{[define PTF]} $G(z)$ is equivalent at the sample times to the
continuous transfer function $G(s)$ driven by ZOH.  This pulse transfer
function can be computed as
%
\begin{equation}
\label{eq:pulsetf}
G(z)=(1-z^{-1})\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\frac{G(s)}{s}}\right\rbrace_{t=kT}}\right\rbrace.
\end{equation}
\end{mylemma}
\red{[Here the dependence on the sampling time T needs to appear somewhere.]}

@@

\noindent where $\sigma_{Q}^{2}$ is the variance of the quantization noise and
$U(z)$ is the z-transform of control signal ({\it i.e.}, the z-transform of
the signal from the digital controller), which is provided to the plant by
means of ZOH.

@@

Eq.~\eqref{eq:deltag_var} indicates that the modelling of the quantization noise
effect on the plant transfer function demands an estimation of the variance
of quantization noise $\sigma_{Q}^{2}$.  To obtain such estimation, the
methodology presented by~\cite{widrow2008quantization}
is used.  The authors provide an algorithm for computing the mean square of
the output noise due to $Q1$ and $Q2$.  Considering that each quantizer
contribution can be represented by a white noise with zero mean
(Assumption~\ref{whitenoise}), the mean square can be considered as the
variance of noise ($\sigma_{Q}^{2}$), which can be calculated as
%
\begin{equation}
\sigma_{Q}^{2}=\frac{\lambda_{1}^{2}\cdot \sum {h_{Q1}^{2}}+\lambda_{2}^{2}\cdot \sum {h_{Q2}^{2}}}{12},
\end{equation}
%
\noindent where $\lambda_{1}$ and $\lambda_{2}$ are the scale factors \red{[define scale factors]} for the
quantization $Q1$ and $Q2$, respectively; and $\sum {h_{Q1}^{2}}$ and $\sum
{h_{Q2}^{2}}$ are the squares sums of the impulse response obtained by means
of a transfer function from $Q1$ and $Q2$ to the plant output, respectively.

@@

These equations are sufficient for computing $\sigma_{Q}^{2}$.  However,
there is still a gap for a complete model for $\Delta{G_{Q}(z)}$ in
Eq.~\eqref{eq:deltag_var}, {\it i.e.}, the z-transform of control action
$U(z)$.  This indicates that the quantization noise effect on a discrete
plant model mainly depends on three main factors:
%
\begin{itemize}
%
\item \textbf{Sample time:} the sample time effect is reflected over pulse transfer function that covers the effects of DAC ant its interpolator (ZOH) \red{[clarify/revise based on the comments made above.]};
%
\item \textbf{ADC and DAC step:} \red{[again, do we need both?]} these are the main component of quantization noise power, but an $8$-bits converter already reduces the effect of quantization noises in the output for negligible values.
%
\item \textbf{Signal-to-noise ratio (SNR): } the dependence on the value of $U(z)$ \red{[notice that now we have reduced the problem to depend on the effect of the reference signal transform R(z) - pls revise.]} shows the importance of the SNR for handling quantization noise; for a high SNR of the digital controller, the quantization effect is negligible. This happens during the transient of the system, when high control actions are employed. However, it can be relevant during the steady state. A digital controller with integral action might be sufficient for eliminating such influence at steady state.
%
\end{itemize}

@@

A more precise description of $\Delta{G_{Q}(z)}$ can be obtained
considering FWL \red{[clarify what you mean with FWL, introduce FWL function explicitly and precisely]} effects on digital controller coefficients, as described in
Bessa {\it et al.}~\cite{Bessa16}, {\it i.e.}, using $FWL[C(z)]$ instead of
the nominal $C(z)$:
%
\begin{equation}
\label{eq:deltaG_final}
\Delta{G_{Q}(z)}=\frac{\sigma^{2}_{Q}}{r}\cdot \frac{1+FWL[C(z)]G(z)}{FWL[C(z)]} \cdot \frac{z-1}{z}.
\end{equation}
